### 2016-05-06 ch02-qlogms/excitation-design.tex
Clean up math + SNR in FD in ch04

In open loop experiments, the scaled average information matrix is given by
\begin{align}
\Fi{\theta} & = \lim_{N\rightarrow\infty}
          \frac{1}{N}
            \sum_{t=1}^{N}
              {\left(\frac{\partial\hat{y}(t)}{\partial\theta}\right)
               \left(\frac{\partial\hat{y}(t)}{\partial\theta}\right)^{\T}},\\
\frac{\partial\hat{y}(t)}{\partial\theta}
    & =\frac{\partial G(s,\theta)}{\partial \theta}u(t).
\end{align}
By Parseval's theorem, $\Fi{\theta}$ can be expressed as
\begin{align}
\Fi{\theta} &= \frac{1}{\pi}
         \int\limits_0^\infty \RealPart
           \left\{      \frac{\partial G(j\omega)}{\partial \theta}
                  \left[\frac{\partial G(j\omega)}{\partial \theta}\right]^{\HT}
           \right\}
           \powerspec{u}(j\omega)
           \dd{\omega} \text{.}
\label{eq:excitation:IF}
\end{align}
The information matrix is a function of the input power spectrum $\powerspec{u}$, which is the only entity that can be used to improve the quality of the estimates. 
A common optimality criterion is D-optimality where $\det{\Fi{\theta}}$ is maximized~\citep{Goodwin1977}.

@@

The information matrix tells us how good the individual parameter estimates are. However, if we are interested in estimating a transfer function $G(s,\theta)$ the quality of this estimate is of more importance. 
It is possible to find an expression for the variance of $\hat{G}(j\omega,\theta)$ using the Gauss approximation formula \citep{Ljung1999, Pintelon2012}. This gives
\begin{align}
\var{\hat{G}(j\omega)} &\approx \frac{\partial G^H(j\omega)}{\partial \theta}\Fi{\theta}^{-1}\frac{\partial G(j\omega)}{\partial \theta}.
\label{eq:excitation:varG}
\end{align}

@@

When $\damping < \frac{1}{\sqrt{2}}$, such a second order system exhibits a peak in the amplitude of the transfer function~\citep{Oppenheim1983}:
  \begin{align}
   \left| G\left(\wM \right) \right| & \isdef
          \max_{\omega} \left| G\left( \omega \right)  \right|
        = \frac{1}
               {2 \damping \sqrt{1-\damping^2}}
               \label{eq:excitation:peakAmplitude}\\
   \wM &
        = \sqrt{1-2\damping^2} \wn
        \label{eq:excitation:peakFrequency}
    \text{.}
  \end{align}
  The $3\unit{dB}$ bandwidth around this resonance peak is given by $\wdB = 2\damping\wn$ for systems that have a sufficiently low damping.

Since information is local in the frequency domain, the input signal power inside $\wdB$, contributes significantly to the information matrix while input power outside $\wdB$ has a smaller contribution. This can be seen by considering the optimal design for a second order system. From \citep[Example 6.4.5]{Goodwin1977} we have that the D-optimal design for \eqref{eq:excitation:2ndOrderSys} is a single sinusoid at frequency
\begin{align}
\omega^\star =\frac%
{\wn
  \sqrt{
    \left(1-2\damping^2\right)
    +
    \sqrt{\left(1-2\damping^2\right)^2+15}
    }
  }
{\sqrt{5}}.
\label{eq:excitation:optFreq}
\end{align}
For lightly-damped systems, i.e $\damping \ll 1$ , $\omega^\star\approx \wn$, which is inside the bandwidth of the system. 
From \eqref{eq:excitation:optFreq} it is clear that the optimal design depends on the resonance of the system to be identified. 
Input design can be based on a guess of $\wn$, however if this guess is wrong, the resulting estimates may be unacceptable~\citep{Rojas2007}.

@@

The information matrix for systems of the type \eqref{eq:excitation:sumSys} will have a block structure commensurate with the dimension of $\theta_i$ given by
\begin{align}
\Fi{\theta} &= \frac{1}{\pi}
\begin{bmatrix}
  \partial G_{1,1}        &\partial G_{1,2}       &\deemph{\cdots}         &\partial G_{1,K}\\
  \partial G_{2,1}        &\partial G_{2,2}       &\deemph{\cdots}         &\partial G_{2,K}\\
  \deemph{\vdots}                  &\deemph{\vdots}                 &\deemph{\ddots}         &\deemph{\vdots}          \\
  \partial G_{K,1}        &\partial G_{K,2}       &\deemph{\cdots}         &\partial G_{K,K}
\end{bmatrix},\\
\partial G_{i,j} &= \int_0^\infty
                      \RealPart\left\{
                          \frac{\partial G_i}
                               {\partial \theta_i}
                          \frac{\partial G_j^{\HT}}
                               {\partial \theta_j}
                                \right\}
                      \phi_u(\omega)\dd\omega.
\end{align}

@@

%FIXME: clean up latex math here, e.g. \frac{\partial ...}{\partial} -> \partialDerivative{...}{...}; use \abs{}; ...

The $(1,1)$ element of the $\partial G_{1,1}$ entry of the information matrix is
\begin{align}
  \int_0^\infty\RealPart\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_1^{\HT}}{\partial \wn[1]}\right\}\frac{1}{\omega}\dd\omega
\end{align}
which evaluates to
\begin{equation}
\frac{\pi }
          {16  \sqrt{1 - \damping^2}\wn^{2} \damping^{3}} 
+ 
\frac{ 2\arctan  
                               \left(
                                       \frac{(2  \damping^{2} - 1) \sqrt{1-\damping^{2}}}
                                                {2  (\damping^{3} - \damping)}
                              \right)}
         {16  \sqrt{1-\damping^{2}} \wn^{2}  \damping^{3}}
+
\frac{{\left(2  \damping^{2} + 1 \right)} }
         {4  \sqrt{1-\damping^{2}} \wn^{2} \damping^{2}}
\label{eq:excitation:auto:full}
\end{equation}
after lengthy calculations.
For $\damping \ll 1$ this can be simplified to
\begin{equation}
\label{eq:excitation:auto}
\int_0^\infty\RealPart\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_1^{\HT}}{\partial \wn[1]}\right\}\frac{1}{\omega}\dd\omega
 = 
 \frac{\pi}
          {8\wn[1]^2}\damping^{-3} 
+ \bigO{\damping^{-2}}
\end{equation}
since the $\arctan$ in \eqref{eq:excitation:auto:full} tends to $\frac{\pi}{2}$ for $\damping \to 0$.
For the corresponding element in $\partial G_{1,2}$ we start by bounding the absolute value of the integral by
\begin{gather}
\left|\int_0^\infty\RealPart\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right|
\qquad \qquad \qquad \qquad \qquad
\label{eq:excitation:cross1} \\
\leq
\int_0^\infty\left|\RealPart\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega
\leq \\
\qquad \qquad \qquad \qquad \qquad
\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega \text{.}\label{eq:excitation:cross}
\end{gather}
To proceed, consider
\begin{align}
\frac{\partial G_1}{\partial \wn[1]} 
  &= \frac{2\damping j\omega\left[1+\frac{j\omega}{\wn[1]\damping}\right]}
  {\wn[1]^2\left[1+2\frac{\damping j\omega}{\wn[1]} + \left(\frac{j\omega}{\wn[1]}\right)^2\right]^2},
\end{align}
which has zeros in $j\omega=\left\{0, \damping\wn[1]\right\}$. 
Hence $\left|\frac{\partial G_1}{\partial \wn[1]} \right|$ has asymptotic slope 20\unit{dB/\decade} for low frequencies, at $j\omega=\damping\wn[1]$ the slope increases to 40\unit{dB/\decade}. 
There are four poles which give a $-40$\unit{dB/\decade} roll-off after the peak. 
The shape of $\frac{\partial G_1}{\partial \wn[1]} $ is illustrated in \figref{fig:excitation:derivatives}.

A crude, albeit useful, bound on the integrand in \eqref{eq:excitation:cross} is
\begin{align}
&f\left(\omega,\theta_1,\theta_2\right)=
  \begin{cases}
   \frac{\omega^2}{\wn[1]^2}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega \leq \wn[1]$,}\\
   \phantom{\frac{\omega^2}{\omega^2}}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\wn[1] < \omega < \wn[2]$,}\\
   \frac{\wn[2]^2}{\omega^2}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega \geq \wn[2]$,}
  \end{cases}
\end{align}
where $\partial G_1^\infty = \left\|\frac{\partial G_1}{\partial \wn[1]}\right\|_\infty$. 
Again, direct calculations give
\begin{align}
\int_0^\infty f\left(\omega,\theta_1,\theta_2\right)\dd\omega = \frac{\wn[1]\left(2-\ln{\frac{\wn[1]^2}{\wn[2]^2}}\right)}{\wn[2]^2}\damping^{-2} + \mathcal{O}(1).
\end{align}
Hence \eqref{eq:excitation:cross1} cannot grow faster than $\mathcal{O}(\damping^{-2})$ as $\damping\rightarrow0$. From this analysis we conclude that:
\begin{itemize}
\item the integral \eqref{eq:excitation:cross1} decreases as $\wn[2]-\wn[1]$ increases, and
\item lower damping further decreases the integral \eqref{eq:excitation:cross1} in comparison to \eqref{eq:excitation:auto}.
\end{itemize}

@@

Based on the above arguments, we approximate the information matrix for lightly-damped systems as a block-diagonal matrix:
\begin{align}
\Fi{\theta}&\approx \begin{bmatrix}\partial G_{1,1} &\deemph{0}\\
                        \deemph{0}&\partial G_{2,2}\end{bmatrix}.
\end{align}
Hence the variance of the estimate $\hat{G}(s,\theta)$ becomes
\begin{align}
\var{\hat{G}(\omega)} &\approx \frac{\partial G^{\HT}(e^{j\omega})}{\partial \theta}\begin{bmatrix}\partial G_{1,1} &\deemph{0}\\
                        \deemph{0}&\partial G_{2,2}\end{bmatrix}^{-1}\frac{\partial G(e^{j\omega})}{\partial \theta}\\
&=\var{\hat{G}_1(\omega)} + \var{\hat{G}_2(\omega)}.
\end{align}

@@

  The choice of the damping is more important.
  Instead of choosing the damping directly, we choose the peak gain $\left| G\left( \wM \right) \right| \in \set{5 \unit{dB}, 10 \unit{dB}, 20 \unit{dB}, 40 \unit{dB} }$ of the transfer function \eqref{eq:excitation:peakAmplitude} to study the effect of differently damped systems.
  However, do remember that $\abs{G(\wM)} = (2\damping\sqrt{1-\damping^2})^{-1}$ for the simplifications introduced above as in \eqref{eq:excitation:peakAmplitude}.

@@

  We then calculate the theoretical variance on the magnitude of the peak in the \gls{FRF} ($\left| G\left( \wM \right) \right|$) as shown in \eqref{eq:excitation:varPeak} for a coarse grid.
  A coarse grid was chosen with $\tilde{\alpha} = 10$; this means that the spacing between the excited lines around the resonance are much larger than the system bandwidth.
  It is obvious that such an excitation signal will not perform well in most cases, as the resonance peak is very likely to be overlooked in such an \gls{FRF}.

>>>>

In open loop experiments, the scaled average information matrix is given by
\begin{align}
\Fi{\theta} & = \lim_{N\rightarrow\infty}
          \frac{1}{N}
            \sum_{t=1}^{N}
              {\left(\PartialDerivative{\hat{y}(t)}{\theta}\right)
               \left(\PartialDerivative{\hat{y}(t)}{\theta}\right)^{\T}},\\
\PartialDerivative{\hat{y}(t)}{\theta}
    & =\PartialDerivative{G(s,\theta)}{\theta}u(t).
\end{align}
By Parseval's theorem, $\Fi{\theta}$ can be expressed as
\begin{align}
\Fi{\theta} &= \frac{1}{\pi}
         \int\limits_0^\infty
        \real{
                  \PartialDerivative{G(j\omega)}{\theta}
                  \left(\PartialDerivative{G(j\omega)}{\theta}\right)^{\HT}
         }
           \powerspec{u}(j\omega)
           \dd{\omega} \text{.}
\label{eq:excitation:IF}
\end{align}
The information matrix is a function of the input power spectrum $\powerspec{u}$, which is the only entity that can be used to improve the quality of the estimates. 
A common optimality criterion is D-optimality where $\det{\Fi{\theta}}$ is maximized~\citep{Goodwin1977}.

@@

The information matrix tells us how good the individual parameter estimates are. However, if we are interested in estimating a transfer function $G(s,\theta)$ the quality of this estimate is of more importance. 
It is possible to find an expression for the variance of $\hat{G}(j\omega,\theta)$ using the Gauss approximation formula \citep{Ljung1999, Pintelon2012}. This gives
\begin{align}
\var{\hat{G}(j\omega)} &
  \approx 
  \left(\PartialDerivative{G(j\omega)}{\theta}\right)^{\HT}
  \Fi{\theta}^{-1}
  \PartialDerivative{G(j\omega)}{\theta}
\label{eq:excitation:varG}
\text{.}
\end{align}

@@

When $\damping < \frac{1}{\sqrt{2}}$, such a second order system exhibits a peak in the amplitude of the transfer function~\citep{Oppenheim1983}:
  \begin{align}
   \abs{G\left(\wM \right) }
   & \isdef
          \max_{\omega} 
            \abs{G\left( \omega \right)}  
        = \frac{1}
               {2 \damping \sqrt{1-\damping^2}}
               \label{eq:excitation:peakAmplitude}\\
   \wM &
        = \sqrt{1-2\damping^2} \wn
        \label{eq:excitation:peakFrequency}
    \text{.}
  \end{align}
  The $3\unit{dB}$ bandwidth around this resonance peak is given by $\wdB = 2\damping\wn$ for systems that have a sufficiently low damping.

Since information is local in the frequency domain, the input signal power inside $\wdB$, contributes significantly to the information matrix while input power outside $\wdB$ has a smaller contribution. This can be seen by considering the optimal design for a second order system. From \citep[Example 6.4.5]{Goodwin1977} we have that the D-optimal design for \eqref{eq:excitation:2ndOrderSys} is a single sinusoid at frequency
\begin{align}
\omega^\star =
\frac{\wn   \sqrt{     \left(1-2\damping^2\right)      +     \sqrt{\left(1-2\damping^2\right)^2+15}    }}
{\sqrt{5}}
\label{eq:excitation:optFreq}
\text{.}
\end{align}
For lightly-damped systems, i.e $\damping \ll 1$ , $\omega^\star\approx \wn$, which is inside the bandwidth of the system. 
From \eqref{eq:excitation:optFreq} it is clear that the optimal design depends on the resonance of the system to be identified. 
Input design can be based on a guess of $\wn$, however if this guess is wrong, the resulting estimates may be unacceptable~\citep{Rojas2007}.

@@

The information matrix for systems of the type \eqref{eq:excitation:sumSys} will have a block structure commensurate with the dimension of $\theta_i$ given by
\begin{align}
\Fi{\theta} &= \frac{1}{\pi}
\begin{bmatrix}
  \partial G_{1,1}        &\partial G_{1,2}       &\deemph{\cdots}         &\partial G_{1,K}\\
  \partial G_{2,1}        &\partial G_{2,2}       &\deemph{\cdots}         &\partial G_{2,K}\\
  \deemph{\vdots}                  &\deemph{\vdots}                 &\deemph{\ddots}         &\deemph{\vdots}          \\
  \partial G_{K,1}        &\partial G_{K,2}       &\deemph{\cdots}         &\partial G_{K,K}
\end{bmatrix},\\
\partial G_{i,j} &= \int_0^\infty
                      \RealPart\left\{
                          \PartialDerivative{G_i}
                               {\theta_i}
                          \PartialDerivative{G_j^{\HT}}
                               {\theta_j}
                                \right\}
                      \phi_u(\omega)\dd\omega.
\end{align}

@@

The $(1,1)$ element of the $\partial G_{1,1}$ entry of the information matrix is
\begin{align}
  \int_0^\infty
    \real{
         \PartialDerivative{G_1}{\wn[1]}
         \cdot
         \PartialDerivative{G_1^{\HT}}{\wn[1]}
    }
      \frac{1}{\omega}
      \dd\omega
\end{align}
which evaluates to
\begin{equation}
\frac{\pi }
          {16  \sqrt{1 - \damping^2}\wn^{2} \damping^{3}} 
+ 
\frac{ 2\arctan  
                               \left(
                                       \frac{(2  \damping^{2} - 1) \sqrt{1-\damping^{2}}}
                                                {2  (\damping^{3} - \damping)}
                              \right)}
         {16  \sqrt{1-\damping^{2}} \wn^{2}  \damping^{3}}
+
\frac{{\left(2  \damping^{2} + 1 \right)} }
         {4  \sqrt{1-\damping^{2}} \wn^{2} \damping^{2}}
\label{eq:excitation:auto:full}
\end{equation}
after lengthy calculations.
For $\damping \ll 1$ this can be simplified to
\begin{equation}
\label{eq:excitation:auto}
\int_0^\infty
\real{
   \PartialDerivative{G_1}{\wn[1]}
   \cdot
   \PartialDerivative{G_1^{\HT}}{\wn[1]}
   }
\frac{1}{\omega}
\dd\omega
 = 
 \frac{\pi}
          {8\wn[1]^2}\damping^{-3} 
+ \bigO{\damping^{-2}}
\end{equation}
since the $\arctan$ in \eqref{eq:excitation:auto:full} tends to $\frac{\pi}{2}$ for $\damping \to 0$.
For the corresponding element in $\partial G_{1,2}$ we start by bounding the absolute value of the integral by
\begin{gather}
\abs{
    \int_0^\infty
    \real{
       \PartialDerivative{G_1}{\wn[1]}
       \cdot
       \PartialDerivative{G_2^{\HT}}{\wn[2]}
    }
    \frac{1}{\omega}
    \dd\omega
}
\qquad \qquad \qquad \qquad \qquad
\label{eq:excitation:cross1} \\
\leq
\int_0^\infty
\abs{
  \real{
    \PartialDerivative{G_1}{\wn[1]}
    \cdot
    \PartialDerivative{G_2^{\HT}}{\wn[2]}
  }
}
\frac{1}{\omega}
\dd\omega
\leq \\
\qquad \qquad \qquad \qquad \qquad
\int_0^\infty
\abs{
  \PartialDerivative{G_1}{\wn[1]}\PartialDerivative{G_2^{\HT}}{\wn[2]}
}
\frac{1}{\omega}
\dd\omega
\label{eq:excitation:cross}
\text{.}
\end{gather}
To proceed, consider
\begin{align}
\PartialDerivative{G_1}{\wn[1]} 
  &= 
     \frac{2\damping j\omega\left[1+\frac{j\omega}{\wn[1]\damping}\right]}
               {\wn[1]^2\left[1+2\frac{\damping j\omega}{\wn[1]} + \left(\frac{j\omega}{\wn[1]}\right)^2\right]^2},
\end{align}
which has zeros in $j\omega=\Set{0, \damping\wn[1]}$. 
Hence $\abs{\PartialDerivative{G_1}{\wn[1]} }$ has asymptotic slope 20\unit{dB/\decade} for low frequencies, at $j\omega=\damping\wn[1]$ the slope increases to 40\unit{dB/\decade}. 
There are four poles which give a $-40$\unit{dB/\decade} roll-off after the peak. 
The shape of $\PartialDerivative{G_1}{\wn[1]} $ is illustrated in \figref{fig:excitation:derivatives}.

A crude, albeit useful, bound on the integrand in \eqref{eq:excitation:cross} is
\begin{align}
&f\left(\omega,\theta_1,\theta_2\right)=
  \begin{cases}
   \frac{\omega^2}{\wn[1]^2}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega \leq \wn[1]$,}\\
   \phantom{\frac{\omega^2}{\omega^2}}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\wn[1] < \omega < \wn[2]$,}\\
   \frac{\wn[2]^2}{\omega^2}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega \geq \wn[2]$,}
  \end{cases}
\end{align}
where $\partial G_1^\infty = \left\|\PartialDerivative{G_1}{\wn[1]}\right\|_\infty$. 
Again, direct calculations give
\begin{align}
\int_0^\infty f\left(\omega,\theta_1,\theta_2\right)\dd\omega = \frac{\wn[1]\left(2-\ln{\frac{\wn[1]^2}{\wn[2]^2}}\right)}{\wn[2]^2}\damping^{-2} + \bigO{1}.
\end{align}
Hence \eqref{eq:excitation:cross1} cannot grow faster than $\bigO{\damping^{-2}}$ as $\damping\rightarrow0$. From this analysis we conclude that:
\begin{itemize}
\item the integral \eqref{eq:excitation:cross1} decreases as $\wn[2]-\wn[1]$ increases, and
\item lower damping further decreases the integral \eqref{eq:excitation:cross1} in comparison to \eqref{eq:excitation:auto}.
\end{itemize}

@@

Based on the above arguments, we approximate the information matrix for lightly-damped systems as a block-diagonal matrix:
\begin{align}
\Fi{\theta}&\approx \begin{bmatrix}\partial G_{1,1} &\deemph{0}\\
                        \deemph{0}&\partial G_{2,2}\end{bmatrix}.
\end{align}
Hence the variance of the estimate $\hat{G}(s,\theta)$ becomes
\begin{align}
\var{\hat{G}(\omega)} &
\approx 
   \PartialDerivative{G^{\HT}(e^{j\omega})}{ \theta}
\begin{bmatrix}
    \partial G_{1,1} &\deemph{0}\\
    \deemph{0}&\partial G_{2,2}
 \end{bmatrix}^{-1}
\PartialDerivative{G(e^{j\omega})}{\theta}\\
&=
     \var{\hat{G}_1(\omega)} 
  + \var{\hat{G}_2(\omega)}
  \text{.}
\end{align}

@@

  The choice of the damping is more important.
  Instead of choosing the damping directly, we choose the peak gain $\abs{ G\left( \wM \right) } \in \set{5 \unit{dB}, 10 \unit{dB}, 20 \unit{dB}, 40 \unit{dB} }$ of the transfer function \eqref{eq:excitation:peakAmplitude} to study the effect of differently damped systems.
  However, do remember that $\abs{G(\wM)} = (2\damping\sqrt{1-\damping^2})^{-1}$ for the simplifications introduced above as in \eqref{eq:excitation:peakAmplitude}.

@@

  We then calculate the theoretical variance on the magnitude of the peak in the \gls{FRF} ($\abs{ G\left( \wM \right) }$) as shown in \eqref{eq:excitation:varPeak} for a coarse grid.
  A coarse grid was chosen with $\tilde{\alpha} = 10$; this means that the spacing between the excited lines around the resonance are much larger than the system bandwidth.
  It is obvious that such an excitation signal will not perform well in most cases, as the resonance peak is very likely to be overlooked in such an \gls{FRF}.
